---
title: "SULF2 is a serological biomarker in MASLD"
author: "Christian Sommerauer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    highlight: tango
subtitle: mRNA correlations of secretet DEGs with fibrosis and NAS in the Govaere
  et al cohort. identify SULF2 as best candidate
---

```{r setup, include=FALSE}
# set path
path <- strsplit(dirname(rstudioapi::getActiveDocumentContext()$path), '/')[[1]]
path <- path[-c((length(path)-0):length(path))]
path <- paste(path, collapse = '/')
knitr::opts_knit$set(root.dir = path)
knitr::opts_chunk$set(echo = T, message = F, warning = F, dev = 'png')
```

# Load annotations, the list of 11 candidate genes, and the tpm-normalized counts from the Govaere et al dataset.
### This analysis cannot be done with the Kozumi dataset as the information about individuals are lacking.
```{r}
getwd()

library(tidyverse)
library(ggpubr)

gene_ann <- read.delim("data/Annotations/230918_february_2023_GeneID_Symbol_Hsap.txt")
GOI.fullTable <- read.delim("results/secreted_overlapping_1e_neg10_03785_koz_gov.txt")
GOI <- GOI.fullTable %>% dplyr::pull("symbol")

govaere = list(tpm_counts = readRDS("data/GSE135251/Processed/Govaere_processed.rds")[[2]] %>% 
                                  tibble::rownames_to_column("GeneID") %>% 
                                  inner_join(gene_ann, .) %>% 
                                  dplyr::select(-GeneID),
                      meta = read.delim("data/GSE135251/Processed/Govaere_meta.txt"))
```

# Plot all pearson correlations as individual plots 
```{r}

output_dir <- "results/mRNA_correlation_plots"
dir.create(path = output_dir, showWarnings = FALSE, recursive = TRUE)

output_dir <- "results/TPM_boxplots"
dir.create(path = output_dir, showWarnings = FALSE, recursive = TRUE)

# Make an empty list
govaere_SULF2_correlation <- list()
# Loop over each candidate gene (GOI)
for (i in 1:length(GOI)) {
gene <- GOI[i]

# Arrange each dataframe (one per candidate gene) in the list in a way that all TPM values, as well as the other parameters like fibrosis, NAS, etc, are in separate columns.
govaere_SULF2_correlation[[i]] <- govaere$tpm_counts[govaere$tpm_counts$symbol==gene,] %>% t() %>% as.data.frame() %>% 
  tibble::rownames_to_column("Run") %>%
  slice(2:217) %>% 
  inner_join(., govaere$meta, by="Run") 

colnames(govaere_SULF2_correlation[[i]]) <- c("Run", "TPM", "STAGE", "Fibrosis_stage", "nas_score", "Cohort", "AT_RISK")

# TPM is as character, has to be changed to numeric.
govaere_SULF2_correlation[[i]]$TPM <-  as.numeric(govaere_SULF2_correlation[[i]]$TPM) 

#Correlations between NAS/Fibrosis stage and SULF2, pearson.Figure 2C-D

ggplot(data = govaere_SULF2_correlation[[i]], aes(x = Fibrosis_stage, y = TPM)) +
  geom_point(shape=21, color='black', aes(fill=AT_RISK), size=3) +
  geom_smooth(method = "lm", color="#C70039") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("#212f3c", "#40E0D0")) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  ylab(paste(gene, "(TPM-norm. gene expression)")) +
  xlab("Fibrosis stage")
ggsave(paste0("results/mRNA_correlation_plots/Fibrosis_cor_Govaere_pearson_",gene, ".pdf"), width=7, height=4)

ggplot(data = govaere_SULF2_correlation[[i]], aes(x = nas_score, y = TPM)) + 
  geom_point(shape=21, color='black', aes(fill=AT_RISK), size=3) +
  geom_smooth(method = "lm", color="#C70039") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("#212f3c", "#40E0D0")) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8), limits=c(0,8)) +
  ylab(paste(gene, "(TPM-norm. gene expression)")) +
  xlab("NAS score")
ggsave(paste0("results/mRNA_correlation_plots/NAS_cor_Govaere_pearson_",gene, ".pdf"), width=7, height=4)

# Supplementary Figure 2C-D (Only SULF2 is shown in the manuscript)
# Expression boxplots
## Fibrosis
ggplot(data = govaere_SULF2_correlation[[i]], aes(x = as.factor(Fibrosis_stage), y = TPM)) + 
  geom_boxplot() +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  ylab(paste(gene, "(TPM-norm. gene expression)")) +
  xlab("Fibrosis stage")
ggsave(paste0("results/TPM_boxplots/TPM_fib_Govaere_",gene, ".pdf"), width=5, height=4)

## NAS
ggplot(data = govaere_SULF2_correlation[[i]], aes(x = as.factor(nas_score), y = TPM)) + 
  geom_boxplot() +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  ylab(paste(gene, "(TPM-norm. gene expression)")) +
  xlab("NAS score")
ggsave(paste0("results/TPM_boxplots/TPM_NAS_Govaere_",gene, ".pdf"), width=5, height=4)

}
```

# This section creates a dataframe that contains all correlations to plot.
```{r}
# Create empty lists.
govaere_gene_list <- list()
govaere_SULF2_correlation_pearson_NAS <- list()
govaere_SULF2_correlation_pearson_fib <- list()

# Like in the section above, loop over the 11 candidate genes, arranging the dataframes columnwise
for (i in 1:length(GOI)) {
gene <- GOI[i]
govaere_gene_list[[i]] <- govaere$tpm_counts[govaere$tpm_counts$symbol==gene,] %>% t() %>% as.data.frame() %>% 
  tibble::rownames_to_column("Run") %>%
  slice(2:217) %>% 
  inner_join(., govaere$meta, by="Run") 

colnames(govaere_gene_list[[i]]) <- c("Run", "TPM", "STAGE", "Fibrosis_stage", "nas_score", "Cohort", "AT_RISK")

# Change TPM from character to numeric
govaere_gene_list[[i]]$TPM <- as.numeric(govaere_gene_list[[i]]$TPM) 

# Add the values for Pearson corr, SULF2 ~ Fibrosis
govaere_SULF2_correlation_pearson_fib[[i]] <- cor(x=govaere_gene_list[[i]]$TPM, y=govaere_gene_list[[i]]$Fibrosis_stage, method = "pearson")
names(govaere_SULF2_correlation_pearson_fib[[i]]) <- GOI[i]
# Add the values for Pearson corr, SULF2 ~ NAS
govaere_SULF2_correlation_pearson_NAS[[i]] <- cor(x=govaere_gene_list[[i]]$TPM, y=govaere_gene_list[[i]]$nas_score, method = "pearson")
names(govaere_SULF2_correlation_pearson_NAS[[i]]) <- GOI[i]
}

# Create a dataframe with correlation values for Pearson corr, SULF2 ~ Fibrosis
govaere_SULF2_correlation_pearson_fib.df <- data.frame(corr_value = unlist(govaere_SULF2_correlation_pearson_fib)) %>% mutate(corr = "Fibrosis") %>% mutate(corr.test = "Pearson") %>% tibble::rownames_to_column("GeneSymbol")
# Create a dataframe with correlation values for Pearson corr, SULF2 ~ NAS
govaere_SULF2_correlation_pearson_NAS.df <- data.frame(corr_value = unlist(govaere_SULF2_correlation_pearson_NAS)) %>% mutate(corr = "NAS") %>% mutate(corr.test = "Pearson") %>% tibble::rownames_to_column("GeneSymbol")

# rbind these dataframes together.
all_correlations <- rbind(govaere_SULF2_correlation_pearson_fib.df, 
                          govaere_SULF2_correlation_pearson_NAS.df) 
```

```{r}
library(patchwork)

# Generate a summary correlation for Pearson correlation in SULF2 ~ Fibrosis
all_correlations_plot1 <- all_correlations %>% 
  filter(corr == "Fibrosis") %>% 
  filter(corr.test == "Pearson") %>% 
  arrange(desc(corr_value)) 
all_correlations_plot1_order <- all_correlations_plot1 %>% pull("GeneSymbol")

p1 <- ggplot(all_correlations_plot1, aes(x=corr_value, y=factor(GeneSymbol, levels = rev(all_correlations_plot1_order)))) +
    geom_point(shape=21, aes(size=corr_value)) +
    scale_size_continuous(name="r", range=c(1,8), limits=c(0.25,0.62)) +
    theme_bw() +
    scale_x_continuous(limits=c(0.28, 0.62), breaks=seq(0.3,0.6,0.1)) +
    xlab("Pearson correlation (r)") +
    ylab("")

# Generate a summary correlation plot for Pearson correlation in SULF2 ~ NAS. Suppl. Figure 2A-B
all_correlations_plot2 <- all_correlations %>%  filter(corr == "NAS") %>% 
  filter(corr.test == "Pearson") %>% 
  arrange(desc(corr_value)) 
all_correlations_plot2_order <- all_correlations_plot2 %>% pull("GeneSymbol") 

p2 <- ggplot(all_correlations_plot2, aes(x=corr_value, y=factor(GeneSymbol, levels = rev(all_correlations_plot2_order)))) +
    geom_point(shape=21, aes(size=corr_value)) +
   scale_size_continuous(name="r", range=c(1,8), limits=c(0.25,0.62)) +
    theme_bw() +
    scale_x_continuous(limits=c(0.28, 0.62), breaks=seq(0.3,0.6,0.1)) +
    xlab("Pearson correlation (r)") +
    ylab("")

# Generate a summary correlation plot with the average between NAS and Fibrosis, for pearson corr. (Figure 2B)
pearson_correlations_avg_plot3 <- govaere_SULF2_correlation_pearson_fib.df %>%   mutate(avg_r_corr = (govaere_SULF2_correlation_pearson_fib.df$corr_value + govaere_SULF2_correlation_pearson_NAS.df$corr_value)/2) %>% 
  arrange(desc(avg_r_corr)) 
pearson_correlations_avg_plot3_order <- pearson_correlations_avg_plot3 %>% pull("GeneSymbol")


p3 <- ggplot(pearson_correlations_avg_plot3, aes(x=avg_r_corr, y=factor(GeneSymbol, levels = rev(pearson_correlations_avg_plot3_order)))) +
    geom_point(shape=21, aes(size=avg_r_corr)) +
    scale_size_continuous(name="r", range=c(1,8), limits=c(0.25,0.62)) +
    theme_bw() +
    scale_x_continuous(limits=c(0.28, 0.62), breaks=seq(0.3,0.6,0.1)) +
    xlab("Pearson correlation (r)") +
    ylab("") 


# Using patchwork, respective plots together in a pdf.
p1+p2+p3
ggsave("results/mRNA_correlation_plots/Pearson_correlation_summary.pdf", width = 25, height =10, units="cm")
```
# Significances for boxplot in Suppl. Fig 2CD
```{r}
SULF2_TPM <- govaere$tpm_counts %>% filter(symbol == "SULF2") %>% t() %>% as.data.frame() %>% tibble::rownames_to_column("Run")
SULF2_TPM <- SULF2_TPM[2:nrow(SULF2_TPM),]
SULF2_TPM2 <- inner_join(SULF2_TPM, govaere$meta, by="Run") 
SULF2_TPM2$V1 <- as.numeric(SULF2_TPM2$V1)
SULF2_TPM2$Fibrosis_stage <- as.factor(SULF2_TPM2$Fibrosis_stage)
SULF2_TPM2$nas_score <- as.factor(SULF2_TPM2$nas_score)

fibrosis_aov <- aov(SULF2_TPM2$V1 ~ SULF2_TPM2$Fibrosis_stage)
TukeyHSD(fibrosis_aov)

nas_score_aov <- aov(SULF2_TPM2$V1 ~ SULF2_TPM2$nas_score)
TukeyHSD(nas_score_aov)
```
```{r}
sessionInfo()
```
 
 