
---
title: "SULF2 as biomarker in MASLD"
author: "Christian Sommerauer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    highlight: tango
subtitle: FLIS2 serum analysis
---


```{r setup, include=FALSE}
# set path
path <- strsplit(dirname(rstudioapi::getActiveDocumentContext()$path), '/')[[1]]
path <- path[-c((length(path)-1):length(path))]
path <- paste(path, collapse = '/')
knitr::opts_knit$set(root.dir = path)
knitr::opts_chunk$set(echo = T, message = F, warning = F, dev = 'png')
```
# Load the packages
```{r}
library(ggplot2)
library(ggpubr)
library(dplyr)
library(viridis)
library(pROC)
```
# Load the data tables
```{r}
FLIS2 <- read.delim("/Volumes/kutterlab/Data/Christian_SULF2_secure/FLIS2/240708_FLIS2_full_table.txt") %>% 
  mutate(BMI = vikt/(height/100*height/100)) %>% # Calculate the BMI from the data.
  mutate(FIB4 = (age * (ASAT_ukat_per_l*60)) / (Thrombocytes * sqrt((ALAT_ukat_per_l*60)))) %>% # microkatal to U -> *60
  mutate(average_conc = (SULF2_rep1 + SULF2_rep2)/2, .before=age) %>% 
  mutate(average_SULF2_ng_ml = average_conc/1000) %>% # for visibility, have SULF2 concentration in ng/ml
  mutate(stage_simple = ifelse(stage=="control", "CTRL", "MASLD")) # generates a new column to simplify the MASLD stages
```

# Box plots to assess MASLD vs control and across different stages.
```{r}
# Plot SULF2 against multiple MASLD stages, fill dots by BMI.
ggplot(FLIS2, aes(y=average_SULF2_ng_ml, x=stage)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=21, aes(fill=BMI), width=0.05, size=2) +
  theme_bw()+
  theme(axis.text = element_text(size=12),
        axis.title = element_text(size=14)) +
  scale_y_continuous(breaks = c(1, 1.5, 2.000, 2.500, 3.000), limits=c(0.600, 3.100)) +
  ylab("SULF2 [ng/ml]") +
  scale_fill_viridis() 
ggsave("MASLD_biomarker_SULF2/results/per_individual_stage_SULF2_FLIS2.pdf", width=4, height=3)
 aov <- aov(data=FLIS2, average_SULF2_ng_ml ~ stage)
TukeyHSD(aov)

# Plot SULF2 in control individuals and patients with MASLD (all stages collapsed)
ggplot(FLIS2, aes(y=average_SULF2_ng_ml, x=stage_simple)) +
  geom_boxplot(outlier.shape = NA) +
  theme_bw() +
  geom_jitter(shape=21, width=0.075, aes(fill=stage)) +
  scale_y_continuous(breaks = c(1, 1.5, 2.000, 2.500, 3.000), limits=c(0.600, 3.000)) +
  ylab("SULF2 [ng/ml]") +
    scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) 
ggsave("MASLD_biomarker_SULF2/results/per_binary_stage_SULF2_FLIS2.pdf", width=3.5, height=3)
wilcox.test(data=FLIS2, average_SULF2_ng_ml ~ stage_simple)
```

# Test individual correlations of the FLIS2 patients with individual parameters.
## CAP, corr. 0.35, p=0.03
```{r}
ggplot(data = FLIS2, aes(x = CAP_dB_per_m, y = average_SULF2_ng_ml)) + 
  geom_point(aes(fill=stage), shape=21, color='black', size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
  ylab("SULF2 [ng/ml]") +
  xlab("CAP [dB/m]")
ggsave("MASLD_biomarker_SULF2/results/SULF2_CAP_corrFLIS2.pdf", width=5, height=2.5)
```

## VCTE, corr. 0.32, p=0.04
```{r}
# THere is one extreme outlier in the VCTE measurement, which will be removed. 
FLIS2_VCTE <- FLIS2 %>%  filter(!VCTE_kPa > 60)
ggplot(data = FLIS2_VCTE, aes(x = VCTE_kPa, y = average_SULF2_ng_ml)) + 
  geom_point(aes(fill=stage), shape=21, color='black', size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
  ylab("SULF2 [ng/ml]") +
  xlab("VCTE [kPa]")
ggsave("MASLD_biomarker_SULF2/results/SULF2_VCTE_corrFLIS2.pdf", width=5, height=2.5)
```

## BMI, corr. 0.42, p=0.005
```{r}
ggplot(data = FLIS2, aes(x = BMI, y = average_SULF2_ng_ml)) + 
  geom_point(aes(fill=stage), shape=21, color='black', size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
  ylab("SULF2 [ng/ml]") +
  xlab("BMI [kg/m2]")
ggsave("MASLD_biomarker_SULF2/results/SULF2_BMI_corrFLIS2.pdf", width=5, height=2.5)
```

## ALT, , corr. 0.14, p=0.4 - one outlier, but no correlation
```{r}
ggplot(data = FLIS2, aes(x = ALAT_ukat_per_l, y = average_SULF2_ng_ml)) + 
  geom_point(aes(fill=stage), shape=21, color='black', size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
  ylab("SULF2 [ng/ml]") +
  xlab("ALT ukat/l")
ggsave("MASLD_biomarker_SULF2/results/SULF2_ALT_corrFLIS2.pdf", width=5, height=2.5)
```
## AST, corr. 0.23, p=0.01
```{r}
ggplot(data = FLIS2, aes(x = ASAT_ukat_per_l, y = average_SULF2_ng_ml)) + 
  geom_point(aes(fill=stage), shape=21, color='black', size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
  ylab("SULF2 [ng/ml]") +
  xlab("AST ukat/l")
ggsave("MASLD_biomarker_SULF2/results/SULF2_AST_corrFLIS2.pdf", width=5, height=2.5)
```
## PRO-C3, 0.29, p=0.06
```{r}
ggplot(data = FLIS2, aes(x = PROC3_ng_per_ml, y = average_SULF2_ng_ml)) + 
  geom_point(aes(fill=stage), shape=21, color='black', size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
  ylab("SULF2 [ng/ml]") +
  xlab("PROC3 [ng/ml]")
ggsave("MASLD_biomarker_SULF2/results/SULF2_PROC3_corrFLIS2.pdf", width=5, height=2.5)
```

## No correlation with thrombocytes or leukocytes or hemoglobin
```{r}
ggplot(data = FLIS2, aes(x = Thrombocytes, y = average_SULF2_ng_ml)) + 
  geom_point(aes(fill=stage), shape=21, color='black', size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
  ylab("SULF2 [ng/ml]") +
  xlab("Thrombocytes")

ggplot(data = FLIS2, aes(x = Leukocytes, y = average_SULF2_ng_ml)) + 
  geom_point(aes(fill=stage), shape=21, color='black', size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
  ylab("SULF2 [ng/ml]") +
  xlab("Leukocytes")

ggplot(data = FLIS2, aes(x = Hemoglobin_g_per_l, y = average_SULF2_ng_ml)) + 
  geom_point(aes(fill=stage), shape=21, color='black', size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
  ylab("SULF2 [ng/ml]") +
  xlab("Hemoglobin_g/l")
```
# ROC curves for MASLD, T2D, and Fibrosis stages.
## Plot the ROC curves together - MASLD and T2D
```{r}
# Select the necessary columns. Here: SULF2 concentration, simplified stages, and T2D
roc_FLIS_table <- FLIS2 %>% dplyr::select(average_SULF2_ng_ml, stage_simple, T2D)
table(roc_FLIS_table$T2D)

# CTRL will be renamed to "0", otherwise "1"
roc_FLIS_table$stage_simple <- ifelse(roc_FLIS_table$stage_simple == 'CTRL', 0, 1)

# Calculate the ROC stats for MASLD and T2D using the pROC package.
roc_curve_MASLD <- roc(roc_FLIS_table$stage_simple, roc_FLIS_table$average_SULF2_ng_ml)
roc_curve_T2D <- roc(roc_FLIS_table$T2D, roc_FLIS_table$average_SULF2_ng_ml)

# Make a new dataframe will all ROC statistics, which can then be plotted.
roc_data <- data.frame(
  fpr_MASLD = rev(1-roc_curve_MASLD$specificities),
  tpr_MASLD = rev(roc_curve_MASLD$sensitivities),
  thresholds_MASLD = rev(roc_curve_MASLD$thresholds),
  fpr_T2D = rev(1-roc_curve_T2D$specificities),
  tpr_T2D = rev(roc_curve_T2D$sensitivities),
  thresholds_T2D= rev(roc_curve_T2D$thresholds))

## I would choose a threshold of 1.11 ng/mL at a FPR of 42% and TPR of 97%


# Plot the ROC curves (both in one plot)
ggplot(roc_data) +
  geom_line(aes(x = fpr_MASLD, y = tpr_MASLD), color = "black", size = 1) +
  geom_line(aes(x = fpr_T2D, y = tpr_T2D), color = "brown", size = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "navy") +
  xlim(0, 1) + ylim(0, 1) +
  labs(
    title = paste0(sprintf("AUC MASLD = %.2f", auc(roc_curve_MASLD)),"\n", sprintf("AUC T2D = %.2f", auc(roc_curve_T2D))),
    x = "False Positive Rate",
    y = "True Positive Rate"
  ) +
  theme_classic()

ggsave("MASLD_biomarker_SULF2/results/SULF2_FLIS2_ROC_MASLD_T2D.pdf", width=3.5, height=2.85)
```
# Plot the difference fibrosis comparisons together.
#F0 vs 1/2/3/4; F0/1 vs 2/3/4; and F0-2 vs F3-4
```{r}
# Select the necessary columns. Here: SULF2 concentration, simplified stages, and three different columns for fibrosis comparisons.
## The comparisons are F0 against F1-4; F0/1 against F2-4; F0-2 against F3/4.
roc_FLIS_table_fibrosis <- FLIS2 %>% dplyr::select(average_SULF2_ng_ml, stage_simple, Fibrosis.stage=Fibrosis.stage, Fibrosis.stage_F0vsall=Fibrosis.stage, 
                                                      Fibrosis.stage_F01vs234=Fibrosis.stage,
                                                      Fibrosis.stage_F012vs34=Fibrosis.stage)


# Rename the relevant fibrosis stages to "0" or "1" dependent what the ROC curves are comparing.
roc_FLIS_table_fibrosis$Fibrosis.stage_F0vsall <- ifelse(roc_FLIS_table_fibrosis$stage_simple == 'CTRL', 0, 
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '0', 0, 
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '1', 1,
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '2', 1,
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '3', 1,
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '4', 1, NA)))))) 

roc_FLIS_table_fibrosis$Fibrosis.stage_F01vs234 <- ifelse(roc_FLIS_table_fibrosis$stage_simple == 'CTRL', 0, 
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '0', 0, 
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '1', 0,
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '2', 1,
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '3', 1,
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '4', 1, NA)))))) 

roc_FLIS_table_fibrosis$Fibrosis.stage_F012vs34 <- ifelse(roc_FLIS_table_fibrosis$stage_simple == 'CTRL', 0, 
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '0', 0, 
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '1', 0,
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '2', 0,
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '3', 1,
                                        ifelse(roc_FLIS_table_fibrosis$Fibrosis.stage == '4', 1, NA)))))) 

# Remove all NAs in the fibrosis columns; 31 columns remain.
roc_FLIS_table_fibrosis <- roc_FLIS_table_fibrosis %>% filter(!is.na(Fibrosis.stage_F0vsall))

roc_curve_F0vsall <- roc(roc_FLIS_table_fibrosis$Fibrosis.stage_F0vsall, roc_FLIS_table_fibrosis$average_SULF2_ng_ml)
roc_curve_F01vs234 <- roc(roc_FLIS_table_fibrosis$Fibrosis.stage_F01vs234, roc_FLIS_table_fibrosis$average_SULF2_ng_ml)
roc_curve_F012vs34 <- roc(roc_FLIS_table_fibrosis$Fibrosis.stage_F012vs34, roc_FLIS_table_fibrosis$average_SULF2_ng_ml)

# Make a new dataframe will all ROC statistics, which can then be plotted.
roc_data <- data.frame(
  fpr_F0vsall = rev(1-roc_curve_F0vsall$specificities),
  tpr_F0vsall = rev(roc_curve_F0vsall$sensitivities),
  thresholds_F0vsall = rev(roc_curve_F0vsall$threshold),
    fpr_F01vs234 = rev(1-roc_curve_F01vs234$specificities),
    tpr_F01vs234 = rev(roc_curve_F01vs234$sensitivities),
    thresholds_F01vs234 = rev(roc_curve_F01vs234$threshold),                        
      fpr_F012vs34 = rev(1-roc_curve_F012vs34$specificities),
      tpr_F012vs34 = rev(roc_curve_F012vs34$sensitivities),
      thresholds_F012vs34 = rev(roc_curve_F012vs34$threshold))

# Plot the ROC curves (three curves in one plot)
ggplot(roc_data) +
  geom_line(aes(x = fpr_F0vsall, y = tpr_F0vsall), color = "black", size = 1) +
  geom_line(aes(x = fpr_F01vs234, y = tpr_F01vs234), color = "#7f5945", size = 1, linetype="dashed") +
  geom_line(aes(x = fpr_F012vs34, y = tpr_F012vs34), color = "#c16e42", size = 1, linetype="dotted") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  xlim(0, 1) + ylim(0, 1) +
  labs(
    title =  paste0(sprintf("AUC F0 vs F1-4, full = %.2f", auc(roc_curve_F0vsall)),"\n", 
                    sprintf("AUC F0/1 vs F2-4, dashed = %.2f", auc(roc_curve_F01vs234)), "\n",
                    sprintf("AUC F0-2 vs F3/4, dotted = %.2f", auc(roc_curve_F012vs34))),
    x = "False Positive Rate",
    y = "True Positive Rate") +
  theme_classic()

ggsave("MASLD_biomarker_SULF2/results/SULF2_FLIS2_ROC_fibrosis_complete.pdf", width=3.5, height=3)
```
# Multi-correlation analysis
```{r}
# Load the libraries necessary for these analyses.
library(corrplot)
library(GGally)

FLIS2$VCTE_kPa[FLIS2$VCTE_kPa > 50] = NA # remove the VCTE outlier again by setting the value NA (this patient is not removed for other correlations)
FLIS2.corr <- FLIS2 
df.flis2 <- data.frame(
  Group = FLIS2.corr$stage,
  SULF2 = FLIS2.corr$average_SULF2_ng_ml,
  BMI = FLIS2.corr$BMI,
  Weight = FLIS2.corr$vikt,
  CAP = FLIS2.corr$CAP_dB_per_m,
  VCTE = FLIS2.corr$VCTE_kPa,
  AST = FLIS2.corr$ASAT_ukat_per_l,
  ALT = FLIS2.corr$ALAT_ukat_per_l)

# MULTI-CORRELATION PLOT
## Set a function to add a correlation line in all scatter plots of the multi-corr plot.
add_corr_line <- function(data, mapping, ...) {
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point(alpha = 0.5, aes(color = Group), size=3) +
    geom_smooth(method = "lm", color = "#252990", ...)
  p
}
## Set a function to modify the upper half of the plot: write "r" instead of "Corr".
custom_upper <- function(data, mapping, ...) {
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  corr_test <- cor.test(x, y)
  corr <- corr_test$estimate
  p_value <- corr_test$p.value
  
  # Determine significance stars
  sig_stars <- if (p_value < 0.001) {
    "***"
  } else if (p_value < 0.01) {
    "**"
  } else if (p_value < 0.05) {
    "*"
  } else {
    ""
  }

  
label <- as.expression(bquote(italic(r) == .(round(corr, 2)) * .(sig_stars)))
 
  ggally_text(label = label,  mapping = aes(x = 0.5, y = 0.5), size = 5, color = "black") +
    theme_void()
}

## Generate the multi-correlation plot.
flis2.p <- ggpairs(df.flis2, 
                   lower = list(continuous = add_corr_line), 
                    upper = list(continuous = custom_upper), ggplot2::aes(color=Group)) +
  theme(
   panel.grid.minor = element_blank(),
    panel.grid.major= element_blank(),
    panel.background = element_rect(fill = "#e5eff5"),
    strip.background = element_rect(fill = "#43566f", color = "white"),
    strip.text = element_text(color = "white", face = "bold") 
  ) 

## changing the color scale using a workaround from stackoverflow. There's no straight forward way implemented by the package. One has to loop over each small plot.
for(i in 1:flis2.p$nrow) {
  for(j in 1:flis2.p$ncol){
    flis2.p[i,j] <- flis2.p[i,j] + 
        scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
        scale_color_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) 
  }
}

flis2.p

ggsave("MASLD_biomarker_SULF2/results/cor_everything_FLIS2.pdf", width=12, height=10)
```

# Supplement. Plot SULF2 levels across genders
```{r}
ggplot(FLIS2, aes(y=average_SULF2_ng_ml, x=gender)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(fill=gender), width=0.05, alpha=0.4, shape=21, size=2) +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_y_continuous(breaks = c(1.000, 1.500, 2.000, 2.500, 3.000), limits=c(0.600, 3.200)) +
  scale_fill_manual(values=c("magenta", "darkblue")) +
  ylab("SULF2 [ng/ml]")

ggsave("MASLD_biomarker_SULF2/results/men_vs_women_SULF2_FLIS2.pdf", width=3.5, height=3)

# Not significantly different between men and women
wilcox.test(data=FLIS2, average_conc ~ gender)
```
# Supplement. Plot SULF2 across age
```{r}
ggplot(data = FLIS2, aes(x = age, y = average_SULF2_ng_ml)) + 
  geom_point(aes(fill=stage), shape=21, color='black', size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
  scale_fill_manual(values=c("#FFC300", "#FF5733", "#C70039", "#581845")) +
   ylab("SULF2 [ng/ml]") 
ggsave("MASLD_biomarker_SULF2/results/241014_age_SULF2_FLIS2.pdf", width=5, height=3)
```
# Cohort characteristics for the table.
```{r}
FLIS2_BMI <- FLIS2 %>% filter(!is.na(BMI)) %>% group_by(stage) %>% 
  mutate(median_BMI = median(BMI)) %>%
  mutate(SD_BMI = sd(BMI)) %>%
  mutate(min_BMI=min(BMI)) %>%
  mutate(max_BMI=max(BMI)) %>%
  dplyr::select(median_BMI, SD_BMI, min_BMI, max_BMI) %>%
  unique()
FLIS2_BMI

FLIS2_age <- FLIS2 %>% group_by(stage) %>% 
  mutate(median_Age = median(age)) %>%
  mutate(SD_Age = sd(age)) %>%
  mutate(min_Age=min(age)) %>%
  mutate(max_Age=max(age)) %>%
  dplyr::select(median_Age, SD_Age, min_Age, max_Age) %>%
  unique()
FLIS2_age

FLIS2_T2D <- FLIS2 %>% group_by(stage) %>% dplyr::select(stage,T2D) %>% table()
FLIS2_T2D <- FLIS2 %>% group_by(stage) %>% dplyr::select(stage,T2D) %>% table()

FLIS2_weight <-  FLIS2 %>% filter(!is.na(vikt)) %>% group_by(stage) %>% 
  mutate(median_weight = median(vikt)) %>%
  mutate(SD_weight = sd(vikt)) %>%
  mutate(min_weight=min(vikt)) %>%
  mutate(max_weight=max(vikt)) %>%
  dplyr::select(median_weight, SD_weight, min_weight, max_weight) %>%
  unique()
FLIS2_ALAT_ukat_per_l

FLIS2_ALAT_ukat_per_l <-  FLIS2 %>% filter(!is.na(ALAT_ukat_per_l)) %>% group_by(stage) %>% 
  mutate(median_ALAT_ukat_per_l = median(ALAT_ukat_per_l)) %>%
  mutate(SD_ALAT_ukat_per_l = sd(ALAT_ukat_per_l)) %>%
  mutate(min_ALAT_ukat_per_l=min(ALAT_ukat_per_l)) %>%
  mutate(max_ALAT_ukat_per_l=max(ALAT_ukat_per_l)) %>%
  dplyr::select(median_ALAT_ukat_per_l, SD_ALAT_ukat_per_l, min_ALAT_ukat_per_l, max_ALAT_ukat_per_l) %>%
  unique()

FLIS2_ASAT_ukat_per_l <-  FLIS2 %>% filter(!is.na(ASAT_ukat_per_l)) %>% group_by(stage) %>% 
  mutate(median_ASAT_ukat_per_l = median(ASAT_ukat_per_l)) %>%
  mutate(SD_ASAT_ukat_per_l = sd(ASAT_ukat_per_l)) %>%
  mutate(min_ASAT_ukat_per_l=min(ASAT_ukat_per_l)) %>%
  mutate(max_ASAT_ukat_per_l=max(ASAT_ukat_per_l)) %>%
  dplyr::select(median_ASAT_ukat_per_l, SD_ASAT_ukat_per_l, min_ASAT_ukat_per_l, max_ASAT_ukat_per_l) %>%
  unique()
FLIS2_ASAT_ukat_per_l

FLIS2_CAP_dB_per_m <-  FLIS2 %>% filter(!is.na(CAP_dB_per_m)) %>% group_by(stage) %>% 
  mutate(median_CAP_dB_per_m = median(CAP_dB_per_m)) %>%
  mutate(SD_CAP_dB_per_m = sd(CAP_dB_per_m)) %>%
  mutate(min_CAP_dB_per_m=min(CAP_dB_per_m)) %>%
  mutate(max_CAP_dB_per_m=max(CAP_dB_per_m)) %>%
  dplyr::select(median_CAP_dB_per_m, SD_CAP_dB_per_m, min_CAP_dB_per_m, max_CAP_dB_per_m) %>%
  unique()
FLIS2_CAP_dB_per_m

FLIS2_VCTE_kPa <-  FLIS2 %>% filter(!is.na(VCTE_kPa)) %>% filter(!VCTE_kPa > 70) %>% group_by(stage) %>% 
  mutate(median_VCTE_kPa = median(VCTE_kPa)) %>%
  mutate(SD_VCTE_kPa = sd(VCTE_kPa)) %>%
  mutate(min_VCTE_kPa=min(VCTE_kPa)) %>%
  mutate(max_VCTE_kPa=max(VCTE_kPa)) %>%
  dplyr::select(median_VCTE_kPa, SD_VCTE_kPa, min_VCTE_kPa, max_VCTE_kPa) %>%
  unique()
FLIS2_VCTE_kPa

FLIS2_SULF2 <-  FLIS2  %>% group_by(stage) %>% 
  mutate(median_SULF2 = median(average_SULF2_ng_ml)) %>%
  mutate(SD_SULF2 = sd(average_SULF2_ng_ml)) %>%
  mutate(min_SULF2=min(average_SULF2_ng_ml)) %>%
  mutate(max_SULF2=max(average_SULF2_ng_ml)) %>%
  dplyr::select(median_SULF2, SD_SULF2, min_SULF2, max_SULF2) %>%
  unique()
FLIS2_SULF2


#Average SULF2 levels
avg_SULF2_ctrl <- FLIS2 %>% filter(stage_simple=="CTRL") %>%
  dplyr::select(stage_simple, average_SULF2_ng_ml)
avg_SULF2_MASLD <- FLIS2 %>% filter(stage_simple=="MASLD") %>%
  dplyr::select(stage_simple, average_SULF2_ng_ml)

"range ctrl"
range(avg_SULF2_ctrl$average_SULF2_ng_ml)
"range MASLD"
range(avg_SULF2_MASLD$average_SULF2_ng_ml)
"median ctrl"
median(avg_SULF2_ctrl$average_SULF2_ng_ml)
"median MASLD"
median(avg_SULF2_MASLD$average_SULF2_ng_ml)
median(avg_SULF2_MASLD$average_SULF2_ng_ml) / median(avg_SULF2_ctrl$average_SULF2_ng_ml)
```
# Safe the resulting table as a tsv.
```{r}
#  Create a function to transpose and remove the stage row
clean_transpose <- function(df) {
  df_t <- t(df)
  # Check if first row is named "stage" and remove it
  if (rownames(df_t)[1] == "stage") {
    df_t <- df_t[-1, ]
  }
  return(df_t)
}


# Apply the function to each dataframe before combining
combined_FLIS2_stats <- as.data.frame(rbind(
  clean_transpose(FLIS2_age),
  clean_transpose(FLIS2_BMI),
  clean_transpose(FLIS2_weight),
  clean_transpose(FLIS2_SULF2),
  clean_transpose(FLIS2_CAP_dB_per_m),
  clean_transpose(FLIS2_VCTE_kPa),
  clean_transpose(FLIS2_ASAT_ukat_per_l),
  clean_transpose(FLIS2_ALAT_ukat_per_l),
  clean_transpose(FLIS2_T2D)
))

positions <- c(33,34)
rownames(combined_FLIS2_stats)[positions] <- c("T2D_no", "T2D_yes")

write.table(combined_FLIS2_stats, "MASLD_biomarker_SULF2/results/FLIS2_cohort_statistics.txt", sep="\t", quote=F)
```

