
---
title: 'SULF2 as biomarker in MASLD'
subtitle: 'Danish cohort serum analysis'
author: 'Christian Sommerauer'
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    highlight: tango
---

```{r setup, include=FALSE}
# set path
path <- strsplit(dirname(rstudioapi::getActiveDocumentContext()$path), '/')[[1]]
path <- path[-c((length(path)-1):length(path))]
path <- paste(path, collapse = '/')
knitr::opts_knit$set(root.dir = path)
knitr::opts_chunk$set(echo = T, message = F, warning = F, dev = 'png')
```

```{r}
library(ggplot2)
library(ggpubr)
library(dplyr)
library(viridis)
```

#Loading the results and making a combined table. The meta data is stored on a access-restricted location
```{r}
batch1_elisa <- read.delim("/Volumes/Kutter/Christian/SULF2_safe_storage/R_import/weight_loss_ELISA1.txt")
batch2_elisa <- read.delim("/Volumes/Kutter/Christian/SULF2_safe_storage/R_import/weight_loss_ELISA2.txt") %>% mutate(study_id = as.character(study_id)) %>% dplyr::select(-rep1, -rep2, -minus_blank_rep1, -minus_blank_rep2, -conc._rep1, -conc._rep2)

meta_full <- read.delim("/Volumes/Kutter/Christian/SULF2_safe_storage/R_import/weight_loss_meta_file.txt") %>%
  mutate(study_id = as.character(study_id)) %>% mutate(gender = as.factor(gender))

combined_table_b1b2 <- rbind(batch1_elisa, batch2_elisa) %>% inner_join(., meta_full, by="study_id") %>% mutate(average_conc=average_conc/1000) %>% mutate(SD_conc=SD_conc/1000) %>% arrange(timepoint,study_id)

combined_table_b1b2_overall_avg <- combined_table_b1b2 %>% 
  group_by(timepoint) %>% mutate(average_SULF2_per_timepoint = mean(average_conc)) %>% dplyr::select(timepoint, average_SULF2_per_timepoint) %>% unique() %>% ungroup() %>% mutate(study_id = 1)
```

#Average SULF2 loss, with range.
```{r}
after_vs_before <- combined_table_b1b2 %>% reframe(V1vsV0 = average_conc[timepoint == "V1"] / average_conc[timepoint == "V0"]) %>%
  mutate(reduction_perc = (1-V1vsV0)*100)

paste('On average, SULF2 serum levels are', (1-round(mean(after_vs_before$V1vsV0),2)),'% lower after weight loss')
paste('The range of SULF2 serum levels are after weight loss range between', round(range(after_vs_before$reduction_perc)[1],2),'%',
      'and', round(range(after_vs_before$reduction_perc)[2],2),'%')

#can also check the geometric mean
paste('On average (geometric mean), SULF2 serum levels are', (1-round(exp(mean(log(after_vs_before$V1vsV0))),2))*100,'% lower after weight loss')
```


# Figure 4A. Generate a plot to show the SULF2 serum levels before and after weight loss in each patient.
```{r}
ggplot(combined_table_b1b2, aes(x=timepoint, group=study_id)) + 
  geom_line(data=combined_table_b1b2, aes(y=average_conc, color=gender), alpha=0.3) + # generates the individual lines for each patient
  geom_point(data=combined_table_b1b2, aes(y=average_conc), size=2, alpha=0.5, color="black") + # generates the individual points for each patient
  geom_line(data=combined_table_b1b2_overall_avg, color="black", aes(y=average_SULF2_per_timepoint), linewidth=1.5) + # thick average line
  geom_point(data=combined_table_b1b2_overall_avg, aes(y=average_SULF2_per_timepoint, group=timepoint), size=2, alpha=0.9, color="black") + # grouping average line to connect it
  theme_bw() +
  scale_color_manual(values=c("darkblue", "magenta")) +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_y_continuous(limits = c(0.300, 2.800), breaks = c(0.500, 1.000, 1.500, 2.000)) +
  coord_cartesian(xlim = c(1.4,1.6)) + # reduce the white space at the sides 
  labs(
    x = "Before_ and after weight loss",
    y = "SULF2 [ng/ml]"
  )

ggsave("MASLD_biomarker_SULF2/results/240617_line_plot_with_mean_line.pdf", width=5, height=3.5)
```

# To generate correlations, we need to append the V0 and V1 values in the same columns.
## To do this, will generate new columns with the same names for V0 and V1, and row bind them
```{r}
combined_table_b1b2_V0 <- combined_table_b1b2 %>% filter(timepoint == "V0") %>% 
  mutate(BMI = bmi_v0) %>% 
  mutate(cholesterol = p_cholesterol_v0) %>% 
  mutate(android_pfat = android_pfat_v0) %>% 
  mutate(weight = weight_v0) %>% 
  mutate(wbtot_pfat = wbtot_pfat_v0) %>%
  mutate(ALT = p_alat_v0) %>%
  mutate(homaIR = homa_ir_v0) %>%
  mutate(metS = ms_z_v1)%>%
  mutate(HDL = p_cholesterol_hdl_v0) %>%
  mutate(LDL = p_cholesterol_ldl_v0) %>%
  mutate(VLDL = p_cholesterol_vldl_v1) %>% 
  mutate(albumin = p_albumin_v0)
  
combined_table_b1b2_V1 <- combined_table_b1b2 %>% filter(timepoint == "V1") %>% 
  mutate(BMI = bmi_v1) %>%  
  mutate(cholesterol = p_cholesterol_v1) %>% 
  mutate(android_pfat = android_pfat_v1) %>% 
  mutate(weight = weight_v1) %>% 
  mutate(wbtot_pfat = wbtot_pfat_v1) %>%
  mutate(ALT = p_alat_v1) %>%
  mutate(homaIR = homa_ir_v1) %>%
  mutate(metS = ms_z_v1) %>%
  mutate(HDL = p_cholesterol_hdl_v1) %>%
  mutate(LDL = p_cholesterol_ldl_v1) %>%
  mutate(VLDL = p_cholesterol_vldl_v1) %>% 
  mutate(albumin = p_albumin_v1)

combined_table_b1b2_corr <- rbind(combined_table_b1b2_V0, combined_table_b1b2_V1)
```

# Statistics for the before and after comparison
```{r}
paired_t_test <- data.frame(study_id = combined_table_b1b2_V0$study_id,
                            difference_V0_V1 = combined_table_b1b2_V0$average_conc - combined_table_b1b2_V1$average_conc,
                            timepoint = as.factor(combined_table_b1b2_V0$timepoint)) %>% 
  mutate(average_difference = mean(difference_V0_V1)) %>% 
  mutate(SD_difference = sd(difference_V0_V1)) 

shapiro.test(paired_t_test$difference_V0_V1) #--> not normally distributed

# calculating the p.values on the differences of after-before or the individual values are both yielding the same results.
t.test(combined_table_b1b2_corr$average_conc ~ combined_table_b1b2_corr$timepoint, paired = TRUE) # Using all values: p-value: 1.186e-08
t.test(paired_t_test$difference_V0_V1) # Directly passing the differences of V0-V1 to the t.test function.  p-value: 1.186e-08

# Since the values are not normally distributed, we however need a non-parametric test
wilcox.test(combined_table_b1b2_corr$average_conc ~ combined_table_b1b2_corr$timepoint, paired = TRUE, alternative = "two.sided") # p-value = 4.075e-10
```
# Invididual correlations between SULF2 serum levels and other parameters.
### Note: the before and after values of the same patients were treated as individual samples for the correlations.

## Figure 4B. BMI
```{r}
ggplot(data = combined_table_b1b2_corr, aes(x = BMI, y = average_conc)) + 
  geom_point(shape=21, color='black', aes(fill=gender), size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("darkblue", "magenta")) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
   ylab("SULF2 [ng/ml]") +
  xlab("BMI")

ggsave("MASLD_biomarker_SULF2/results/240112_WLcoh_corr_plot_BMI.pdf", width=5, height=3)
```

## Figure 4C. ALT
```{r}
ggplot(data = combined_table_b1b2_corr, aes(x = log2(ALT), y = average_conc)) + 
  geom_point(shape=21, color='black', aes(fill=gender), size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("darkblue", "magenta")) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
   ylab("SULF2 [ng/ml]") +
  xlab("ALT (log2)")

ggsave("MASLD_biomarker_SULF2/results/240617_WLcoh_ALT_corr_plot.pdf", width=5, height=3)
```

## Figure 4D. Android fat percentage
```{r}
ggplot(data = combined_table_b1b2_corr, aes(x = android_pfat, y = average_conc)) + 
  geom_point(shape=21, color='black', aes(fill=gender), size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("darkblue", "magenta")) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
   ylab("SULF2 [ng/ml]") +
  xlab("Android pfat")

ggsave("MASLD_biomarker_SULF2/results/240617_WLcoh_android_pfat_corr_plot.pdf", width=5, height=3)
```

## Figure 4E. Cholesterol
```{r}
ggplot(data = combined_table_b1b2_corr, aes(x = cholesterol, y = average_conc)) + 
  geom_point(shape=21, color='black', aes(fill=gender), size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("darkblue", "magenta")) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
   ylab("SULF2 [ng/ml]") +
  xlab("Serum cholesterol")

ggsave("MASLD_biomarker_SULF2/results/240112_WLcoh_corr_plot_cholesterol.pdf", width=5, height=3)
```


# Figure 4F. WB fat percentage
```{r}
ggplot(data = combined_table_b1b2_corr, aes(x = wbtot_pfat, y = average_conc)) + 
  geom_point(shape=21, color='black', aes(fill=gender), size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("darkblue", "magenta")) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
   ylab("SULF2 [ng/ml]") +
  xlab("wb tot")

ggsave("MASLD_biomarker_SULF2/results/240617_WLcoh_wbtot_corr_plot.pdf", width=5, height=3)
```

## Figure 4G. HOMA-IR
```{r}
ggplot(data = combined_table_b1b2_corr, aes(x = homaIR, y = average_conc)) + 
  geom_point(shape=21, color='black', aes(fill=gender), size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("darkblue", "magenta")) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
   ylab("SULF2 [ng/ml]") +
  xlab("HOMA-IR")

ggsave("MASLD_biomarker_SULF2/results/240617_WLcoh_HOMAIR_corr_plot.pdf", width=5, height=3)
```

## Figure 4H. LDL
```{r}
ggplot(data = combined_table_b1b2_corr, aes(x = LDL, y = average_conc)) + 
  geom_point(shape=21, color='black', aes(fill=gender), size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("darkblue", "magenta")) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
   ylab("SULF2 [ng/ml]") +
  xlab("LDL")

ggsave("MASLD_biomarker_SULF2/results/240617_WLcoh_LDL_corr_plot.pdf", width=5, height=3)
```

## Figure 4I. MetS z-score
```{r}
ggplot(data = combined_table_b1b2_corr, aes(x = metS, y = average_conc)) + 
  geom_point(shape=21, color='black', aes(fill=gender), size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("darkblue", "magenta")) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
   ylab("SULF2 [ng/ml]") +
  xlab("MetS")

ggsave("MASLD_biomarker_SULF2/results/240617_WLcoh_metS_corr_plot.pdf", width=5, height=3)
```

# Supplementary Figure 4A. Multi-correlation plot showing individual correlations and scatterplots between different parameters.
```{r}
# Load the libraries
library(corrplot)
library(GGally)


df.slite <- data.frame(
  SULF2 = combined_table_b1b2_corr$average_conc,
  BMI = combined_table_b1b2_corr$BMI,
  Weight = combined_table_b1b2_corr$weight,
  WB.fat = combined_table_b1b2_corr$wbtot_pfat,
  Andr.fat = combined_table_b1b2_corr$android_pfat,
  HOMA.IR = combined_table_b1b2_corr$homaIR,
  Cholesterol = combined_table_b1b2_corr$cholesterol,
  metS = combined_table_b1b2_corr$metS,
  ALT =log2(combined_table_b1b2_corr$ALT))
  

 #Define a custom function to add a correlation line
add_corr_line <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point(alpha = 0.3) +
    geom_smooth(method = "lm", color = "blue", ...)
  p
}

custom_upper <- function(data, mapping, ...) {
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  corr_test <- cor.test(x, y)
  corr <- corr_test$estimate
  p_value <- corr_test$p.value
  
  # Determine significance stars
  sig_stars <- if (p_value < 0.001) {
    "***"
  } else if (p_value < 0.01) {
    "**"
  } else if (p_value < 0.05) {
    "*"
  } else {
    ""
  }

  
label <- as.expression(bquote(italic(r) == .(round(corr, 2)) * .(sig_stars)))
 
  ggally_text(label = label,  mapping = aes(x = 0.5, y = 0.5), size = 4, color = "black") +
    theme_void()
}

slite.p <- ggpairs(df.slite, 
                   lower = list(continuous = add_corr_line), 
                    upper = list(continuous = custom_upper)) +# ggplot2::aes(color=Group)) +
  theme(
   panel.grid.minor = element_blank(),
    panel.grid.major= element_blank(),
    panel.background = element_rect(fill = "#e5eff5"),
    strip.background = element_rect(fill = "#43566f", color = "white"),
    strip.text = element_text(color = "white", face = "bold") 
  ) 

slite.p

ggsave("MASLD_biomarker_SULF2/results/240702_cor_everything_WL.pdf", width=11, height=8)
```
# Supplementary Figure 4B-C. SULF2 across genders and age
```{r}
ggplot(combined_table_b1b2, aes(y=average_conc, x=factor(gender, levels = c(2,1)))) +
  geom_boxplot(outlier.shape = NA) +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  geom_jitter(width=0.05, alpha=0.4) +
  facet_wrap(~timepoint) +
  scale_y_continuous(breaks = c(0, 0.500, 1.000, 1.500, 2.000, 2.500), limits=c(0, 2.800)) +
  ylab("SULF2 [ng/ml]")

ggsave("MASLD_biomarker_SULF2/results/240617_WL_men_vs_women_SULF2_FLIS2.pdf", width=6, height=3)

# Not significantly different between men and women
wilcox.test(data=combined_table_b1b2_V0, average_conc ~ gender)
wilcox.test(data=combined_table_b1b2_V1, average_conc ~ gender)

ggplot(data = combined_table_b1b2, aes(x = age_inclusion, y = average_conc)) + 
  geom_point(shape=21, color='black', aes(fill=gender), size=3) +
  geom_smooth(method = "lm", se = FALSE, color="black") +
  theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title = element_text(size=14)) +
  scale_fill_manual(values=c("darkblue", "magenta")) +
  facet_wrap(~timepoint) +
  stat_cor(method="pearson", p.digits=0, size=4, cor.coef.name = "r") +
   ylab("SULF2 conc. in ng/ml") +
  xlab("Age")

ggsave("MASLD_biomarker_SULF2/results/240617_WLcor_age_corr_plot.pdf", width=7, height=3)
```

# Table 2. Characteristics for table with patient characteristics for the Danish Weight loss cohort.
```{r}
#BMI
combined_table_b1b2_v0_BMI <- combined_table_b1b2 %>% filter(timepoint =="V0") %>% 
  mutate(mean_BMI = mean(bmi_v0)) %>%
  mutate(SD_BMI = sd(bmi_v0)) %>% 
  mutate(min_BMI = min(bmi_v0)) %>%
  mutate(max_BMI = max(bmi_v0)) %>%
  dplyr::select(timepoint, mean_BMI, SD_BMI, min_BMI, max_BMI) %>% unique()
#BMI
combined_table_b1b2_v1_BMI <- combined_table_b1b2 %>% filter(timepoint =="V1") %>% 
  mutate(mean_BMI = mean(bmi_v1)) %>%
  mutate(SD_BMI = sd(bmi_v1)) %>% 
  mutate(min_BMI = min(bmi_v1)) %>%
  mutate(max_BMI = max(bmi_v1)) %>%
  dplyr::select(timepoint, mean_BMI, SD_BMI, min_BMI, max_BMI) %>% unique()
#Age 
combined_table_b1b2_v0_age <- combined_table_b1b2 %>% filter(timepoint =="V0") %>% 
  mutate(mean_age = mean(age_inclusion)) %>%
  mutate(SD_age = sd(age_inclusion)) %>% 
  mutate(min_age = min(age_inclusion)) %>%
  mutate(max_age = max(age_inclusion)) %>%
  dplyr::select(timepoint, mean_age, SD_age, min_age, max_age) %>% unique()
#Weight
combined_table_b1b2_v0_weight <- combined_table_b1b2 %>% filter(timepoint =="V0") %>% 
  mutate(mean_weight = mean(weight_v0)) %>%
  mutate(SD_weight = sd(weight_v0)) %>% 
  mutate(min_weight = min(weight_v0)) %>%
  mutate(max_weight = max(weight_v0)) %>%
  dplyr::select(timepoint, mean_weight, SD_weight, min_weight, max_weight) %>% unique()
#Weight
combined_table_b1b2_v1_weight <- combined_table_b1b2 %>% filter(timepoint =="V1") %>% 
  mutate(mean_weight = mean(weight_v1)) %>%
  mutate(SD_weight = sd(weight_v1)) %>% 
  mutate(min_weight = min(weight_v1)) %>%
  mutate(max_weight = max(weight_v1)) %>%
  dplyr::select(timepoint, mean_weight, SD_weight, min_weight, max_weight) %>% unique()
#total fat percentage
combined_table_b1b2_v0_totfat <- combined_table_b1b2 %>% filter(timepoint =="V0") %>% 
  mutate(mean_totfat = mean(wbtot_pfat_v0)) %>%
  mutate(SD_totfat= sd(wbtot_pfat_v0)) %>% 
  mutate(min_totfat = min(wbtot_pfat_v0)) %>%
  mutate(max_totfat = max(wbtot_pfat_v0)) %>%
  dplyr::select(timepoint, mean_totfat, SD_totfat, min_totfat, max_totfat) %>% unique()
#total fat percentage
combined_table_b1b2_v1_totfat <- combined_table_b1b2 %>% filter(timepoint =="V1") %>% 
  mutate(mean_totfat = mean(wbtot_pfat_v1)) %>%
  mutate(SD_totfat= sd(wbtot_pfat_v1)) %>% 
  mutate(min_totfat = min(wbtot_pfat_v1)) %>%
  mutate(max_totfat = max(wbtot_pfat_v1)) %>%
  dplyr::select(timepoint, mean_totfat, SD_totfat, min_totfat, max_totfat) %>% unique()
#ALT
combined_table_b1b2_v0_ALT <- combined_table_b1b2 %>% filter(timepoint =="V0") %>% 
  mutate(mean_ALT = mean(p_alat_v0)) %>%
  mutate(SD_ALT = sd(p_alat_v0)) %>% 
  mutate(min_ALT = min(p_alat_v0)) %>%
  mutate(max_ALT = max(p_alat_v0)) %>%
  dplyr::select(timepoint, mean_ALT, SD_ALT, min_ALT, max_ALT) %>% unique()
#ALT
combined_table_b1b2_v1_ALT <- combined_table_b1b2 %>% filter(timepoint =="V1") %>% 
  mutate(mean_ALT = mean(p_alat_v1)) %>%
  mutate(SD_ALT = sd(p_alat_v1)) %>% 
  mutate(min_ALT = min(p_alat_v1)) %>%
  mutate(max_ALT = max(p_alat_v1)) %>%
  dplyr::select(timepoint, mean_ALT, SD_ALT, min_ALT, max_ALT) %>% unique()
#Serum cholesterol
combined_table_b1b2_v0_cholesterol <- combined_table_b1b2 %>% filter(timepoint =="V0") %>% 
  mutate(mean_chol = mean(p_cholesterol_v0)) %>%
  mutate(SD_chol = sd(p_cholesterol_v0)) %>% 
  mutate(min_chol = min(p_cholesterol_v0)) %>%
  mutate(max_chol = max(p_cholesterol_v0)) %>%
  dplyr::select(timepoint, mean_chol, SD_chol, min_chol, max_chol) %>% unique()
#Serum cholesterol
combined_table_b1b2_v1_cholesterol <- combined_table_b1b2 %>% filter(timepoint =="V1") %>% 
  mutate(mean_chol = mean(p_cholesterol_v1)) %>%
  mutate(SD_chol = sd(p_cholesterol_v1)) %>% 
  mutate(min_chol = min(p_cholesterol_v1)) %>%
  mutate(max_chol = max(p_cholesterol_v1)) %>%
  dplyr::select(timepoint, mean_chol, SD_chol, min_chol, max_chol) %>% unique()
#Android fat percentage
combined_table_b1b2_v0_android <- combined_table_b1b2 %>% filter(timepoint =="V0") %>% 
  mutate(mean_androi = mean(android_pfat_v0)) %>%
  mutate(SD_androi = sd(android_pfat_v0)) %>% 
  mutate(min_androi = min(android_pfat_v0)) %>%
  mutate(max_androi = max(android_pfat_v0)) %>%
  dplyr::select(timepoint, mean_androi, SD_androi, min_androi, max_androi) %>% unique()
#Android fat percentage
combined_table_b1b2_v1_android <- combined_table_b1b2 %>% filter(timepoint =="V1") %>% 
  mutate(mean_androi = mean(android_pfat_v1)) %>%
  mutate(SD_androi = sd(android_pfat_v1)) %>% 
  mutate(min_androi = min(android_pfat_v1)) %>%
  mutate(max_androi = max(android_pfat_v1)) %>%
  dplyr::select(timepoint, mean_androi, SD_androi, min_androi, max_androi) %>% unique()
#LDL - two patients removed due to NA
combined_table_b1b2_v0_LDL <- combined_table_b1b2 %>% filter(timepoint =="V0") %>% filter(!study_id=="127", !study_id=="172") %>%
  mutate(mean_LDL = mean(p_cholesterol_ldl_v0)) %>%
  mutate(SD_LDL = sd(p_cholesterol_ldl_v0)) %>% 
  mutate(min_LDL = min(p_cholesterol_ldl_v0)) %>%
  mutate(max_LDL = max(p_cholesterol_ldl_v0)) %>%
  dplyr::select(timepoint, mean_LDL, SD_LDL, min_LDL, max_LDL) %>% unique()
#LDL - two patients removed due to NA
combined_table_b1b2_v1_LDL <- combined_table_b1b2  %>% filter(!study_id=="127", !study_id=="172") %>%
  mutate(mean_LDL = mean(p_cholesterol_ldl_v1)) %>%
  mutate(SD_LDL = sd(p_cholesterol_ldl_v1)) %>% 
  mutate(min_LDL = min(p_cholesterol_ldl_v1)) %>%
  mutate(max_LDL = max(p_cholesterol_ldl_v1)) %>%
  dplyr::select(timepoint, mean_LDL, SD_LDL, min_LDL, max_LDL) %>% unique()
#HOMA-IR - removed several patients due to NA
combined_table_b1b2_v0_homa1 <- combined_table_b1b2 %>% filter(timepoint =="V0") 
combined_table_b1b2_v0_homa1 <- combined_table_b1b2_v0_homa1[!is.na(combined_table_b1b2_v0_homa1$homa_ir_v0),]

combined_table_b1b2_v0_homaIR <- combined_table_b1b2_v0_homa1 %>% 
  mutate(mean_IR = mean(homa_ir_v0)) %>%
  mutate(SD_IR = sd(homa_ir_v0)) %>% 
  mutate(min_IR = min(homa_ir_v0)) %>%
  mutate(max_IR = max(homa_ir_v0)) %>%
  dplyr::select(timepoint, mean_IR, SD_IR, min_IR, max_IR) %>% unique()
#HOMA-IR - removed several patients due to NA
combined_table_b1b2_v1_homa1 <- combined_table_b1b2 %>% filter(timepoint =="V1") 
combined_table_b1b2_v1_homa1 <- combined_table_b1b2_v1_homa1[!is.na(combined_table_b1b2_v1_homa1$homa_ir_v1),]

combined_table_b1b2_v1_homaIR <- combined_table_b1b2_v1_homa1 %>%
  mutate(mean_IR = mean(homa_ir_v1)) %>%
  mutate(SD_IR = sd(homa_ir_v1)) %>% 
  mutate(min_IR = min(homa_ir_v1)) %>%
  mutate(max_IR = max(homa_ir_v1)) %>%
  dplyr::select(timepoint, mean_IR, SD_IR, min_IR, max_IR) %>% unique()
#MetS z-score
combined_table_b1b2_v0_mets <- combined_table_b1b2 %>% filter(timepoint =="V0") %>%
  mutate(mean_metS = mean(ms_z_v0)) %>%
  mutate(SD_metS = sd(ms_z_v0)) %>% 
  mutate(min_metS = min(ms_z_v0)) %>%
  mutate(max_metS = max(ms_z_v0)) %>%
  dplyr::select(timepoint, mean_metS, SD_metS, min_metS, max_metS) %>% unique()
#MetS z-score
combined_table_b1b2_v1_mets <- combined_table_b1b2 %>% filter(timepoint =="V1") %>%
  mutate(mean_metS = mean(ms_z_v1)) %>%
  mutate(SD_metS = sd(ms_z_v1)) %>% 
  mutate(min_metS = min(ms_z_v1)) %>%
  mutate(max_metS = max(ms_z_v1)) %>%
  dplyr::select(timepoint, mean_metS, SD_metS, min_metS, max_metS) %>% unique()

final_v0 <- cbind(combined_table_b1b2_v0_BMI, combined_table_b1b2_v0_ALT, combined_table_b1b2_v0_android, combined_table_b1b2_v0_cholesterol, combined_table_b1b2_v0_totfat, combined_table_b1b2_v0_weight, combined_table_b1b2_v0_homaIR, combined_table_b1b2_v0_LDL, combined_table_b1b2_v0_mets, combined_table_b1b2_v0_age)
final_v0 <- final_v0[, !duplicated(colnames(final_v0))]

final_v1 <- cbind(combined_table_b1b2_v1_BMI, combined_table_b1b2_v1_ALT, combined_table_b1b2_v1_android, combined_table_b1b2_v1_cholesterol, combined_table_b1b2_v1_totfat, combined_table_b1b2_v1_weight, combined_table_b1b2_v1_homaIR, combined_table_b1b2_v1_LDL, combined_table_b1b2_v1_mets, combined_table_b1b2_v0_age)
final_v1 <- final_v1[, !duplicated(colnames(final_v1))] %>% unique()

final_table <- rbind(final_v0, final_v1)
final_table
```